---
title: "Cytometry Workflow 1 - Preparing the data"
output: 
    html_document:
        df_print: paged
        code_folding: show
        toc: yes
        toc_float: yes
        toc_depth: 4
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(targets)
# tar_unscript() # this line resets the whole pipeline!
```

# Globals

We first define some global options/functions common to all targets.

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c(c("MASS","cluster",
                              "flowWorkspace","flowCore",
                              "hilbertSimilarity","entropy",
                              "Radviz","cytofan",
                              "lme4","emmeans",
                              "FlowSOM","Rtsne",
                              "yaml","tidyverse")))
source("scripts/900_functions.R")
```

# Parameters

We need to define:

 - the path to the `GatingSet` object to analyze
 - the full node path to the root population of interest
 - the list of markers corresponding to the different categories
     - *reference* markers that are part of the gating scheme and will be negative or positive for all cells
     - *phenotypic* markers that are differentiating populations within the population of interest (bimodal)
     - *functional* markers that are indicative of the status of the cells (continuous)
     - all other markers are considered *CyTOF*

```{targets params}
list(tar_target(params_path,
                file.path("process","workshop","live")),
     tar_target(params_population,
                "/Bead Removal/Intact cells/Singlets/Viable/CD45+/CD45:CD3"),
     tar_target(params_ref_markers,
                paste0(str_replace_all("cells[1-2]|dead|CD45|CD3",
                                       "\\|",
                                       "$|"),
                       "$")),
     tar_target(params_pheno_markers,
                paste0(str_replace_all("CD127|CD4|CD8|TBET|CD25|gdTCR|CD45RA|CD27|FOXP3|CD56",
                                       "\\|",
                                       "$|"),
                       "$")),
     tar_target(params_func_markers,
                paste0(str_replace_all("CCR6|IL4|CD38|IL22|CD28|CXCR4|CCR4|GMCSF|TGFb|IL17a|CXCR3|CD161|IL-10|CCR7|IL21|CTLA4|CXCR5|IL17F|HLA-DR|IFNg|TNF",
                                       "\\|",
                                       "$|"),
                       "$"))
)
```

# Sample annotation

Sample annotation can be either parsed from FCS file names (as below) or loaded from a file; is **must** include at least a `name` column corresponding to the FCS file names as returned by `sampleNames(.gs)` below.

```{targets annots, tar_simple=TRUE}
.gs <- load_gs(params_path)
tibble(name = sampleNames(.gs)) %>% 
    mutate(Experiment = str_extract(name,"[0-9]{6}"),
           Experiment = factor(Experiment),
           Stimulation = str_extract(name,"(rest)|(stim)"),
           Stimulation = factor(Stimulation,
                                levels = c("rest","stim")),
           Individual = str_extract(name,"SK[0-9]{3}"),
           Individual = factor(Individual))
```

# Marker annotation

Marker annotation uses the parameters identified above to identify channels that are important for the analysis; in particular *phenotypic* and *functional* markers will be used throughout. This is also an opportunity to clean up the `desc` field, returning human-readable channel names.

It's easier to get started from the file associated with the `GatingSet` being analyzed as shown below:

```{targets markers, tar_simple=TRUE}
.gs <- load_gs(params_path)
as_tibble(pData(parameters(gs_cyto_data(.gs)[[1]])),
                      rownames = "id") %>% 
    mutate(desc_orig = desc) %>% 
    separate(desc,
             c("isotope","desc"),
             sep = "_") %>% 
    mutate(type = case_when(
        str_detect(desc,params_ref_markers) ~ "reference",
        str_detect(desc,params_pheno_markers) ~ "phenotypic",
        str_detect(desc,params_func_markers) ~ "functional",
        TRUE ~ "CyTOF"),
    desc = if_else(is.na(desc),
                   name,
                   desc),
    desc = make.names(desc),
    type = factor(type,
                  levels = c("reference","phenotypic","functional","CyTOF")))
```

# Extract the data

To run the analysis we will need to access the expression values stored in the FCS files through the `exprs` command:

```{targets gs-to-tibble}
list(
    tar_target(live.mats,{
        .gs <- load_gs(params_path)
        lapply(.gs,
               function(gh) {
                   exprs(gh_pop_get_data(gh,
                                         params_population))})
    }),
    tar_target(live.df,{
        bind_rows(lapply(live.mats,
                         as_tibble),
                  .id = "name")
    })
)
```

# Run the pipeline

```{r}
tar_make()
```

# Quick overview

## Sample annotations

```{r}
tar_read(annots)
```

## Marker annotations

```{r}
tar_read(markers)
```

