---
title: "Cytometry Workflow 2 - Quality Control"
output: 
    html_document:
        df_print: paged
        code_folding: show
        toc: yes
        toc_float: yes
        toc_depth: 4
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(targets)
```

# Globals

We first define some global options/functions common to all targets.

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c(c("MASS","cluster",
                              "flowWorkspace","flowCore",
                              "hilbertSimilarity","entropy",
                              "Radviz","cytofan",
                              "lme4","emmeans",
                              "FlowSOM","Rtsne",
                              "yaml","tidyverse")))
```

# Marker Enrichment Modeling (MEM)

To identify potential batch effects associated to channel intensity we use the **M**arker **E**nrichment **M**odeling method ([Diggins 2017](https://doi.org/10.1038/nmeth.4149)) with some adapations:

```{targets live.mem, tar_simple=TRUE, tar_interactive=FALSE}
.mats.med <- lapply(live.mats,function(mt,ch) {
    mm <- apply(mt[,ch],2,median)
    return(mm)
},with(markers,name[type %in% c("phenotypic","functional")]))

.mats.iqr <- lapply(live.mats,function(mt,ch) {
    mi <- apply(mt[,ch],2,IQR)
    if(min(mi)==0) {
        thresh <- min(mi[mi>0])
        thresh <- 10^floor(log10(thresh))
        mi <- replace(mi,mi==0,thresh)
    }
    return(mi)
},with(markers,name[type %in% c("phenotypic","functional")]))

.ref <- pam(bind_rows(.mats.med),1)
.ref <- .ref$id.med

.med.diffs <- sweep(bind_rows(.mats.med),2,.mats.med[[.ref]],`-`)
.iqr.ratio <- sweep(bind_rows(.mats.iqr),2,.mats.iqr[[.ref]],`/`)
.iqr.ratio <- 1/.iqr.ratio

.mem <- abs(.med.diffs)+.iqr.ratio-1
.mem[.med.diffs<0] <- -.mem[.med.diffs<0]
.mem <- asinh(.mem)
rownames(.mem) <- names(.mats.med)
.mem <- .mem[-.ref,]
.mem
```

We prepare a basic plot for reporting:

```{targets mem-plot, tar_interactive=FALSE}
tar_target(live.mem.plot, {
    as_tibble(live.mem,rownames = "id") %>% 
        gather("name","MEM",-id) %>% 
        left_join(markers[,c("name","desc")],
                  by = "name") %>% 
        left_join(annots,
                  by = c("id"="name")) %>% 
        mutate(desc = factor(desc),
               desc = reorder(desc,abs(MEM),mean),
               desc = factor(desc,
                             levels = rev(levels(desc)))) %>% 
        ggplot(aes(x = desc,
                   y = MEM))+
        geom_line(aes(group = id),
                  color = "grey80")
})
```

# Hilbert Similarity

To identify potential batch effects we compute the similarity between samples using the `hilbertSimilarity` package ([doi:10.5281/zenodo.3557362](https://zenodo.org/record/3557362)). First we define which markers to include in the analysis:

```{targets params-hilbert}
tar_target(params_hilbert_channels,{
    with(markers,
         name[type %in%c("phenotypic","functional")])
})
```

Next we associate every cell in the dataset to a value on the Hilbert curve defined over all selected channels:

```{targets live.hilbert, tar_simple=TRUE, tar_interactive=FALSE}
## create the cuts 
.nbins <- 2

.cuts <- live.df %>% 
    select(all_of(params_hilbert_channels)) %>% 
    as.matrix() %>% 
    make.cut(.,
             n=.nbins+1,
             count.lim=40)

## apply the cuts
.cutMat <- live.df %>% 
    select(all_of(params_hilbert_channels)) %>% 
    as.matrix() %>% 
    do.cut(.,.cuts,
           type='combined')

## generate the hilbert index
do.hilbert(.cutMat,.nbins)
```

Based on the Hilbert index we compute the composition of every sample and determine sample similarity using the Jensen-Shannon distance:

```{targets live.hilbert.dist, tar_simple=TRUE, tar_interactive = FALSE}
.counts <- table(live.df$name,
                 live.hilbert)

js.dist(.counts)
```

To visualize the similarity between samples we create a simple projection using the **M**ulti**D**imensional **S**caling method:

```{targets live.hilbert.proj, tar_simple=TRUE, tar_interactive = FALSE}
.proj <- isoMDS(live.hilbert.dist)

as_tibble(.proj$points,
          .name_repair = "universal",
          rownames = "name") %>% 
    rename(x = `...1`,
           y = `...2`) %>% 
    left_join(annots,
              by = "name") %>% 
    left_join(live.df %>% 
                  count(name,
                        name = "N"),
              by = "name") %>% 
    ggplot(aes(x = x,
               y = y))
```

# Run the pipeline

```{r}
tar_make()
```
