---
title: "Cytometry Workflow 3 - Visualizations"
output: 
    html_document:
        df_print: paged
        code_folding: show
        toc: yes
        toc_float: yes
        toc_depth: 4
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(targets)
```

# Globals

We first define some global options/functions common to all targets.

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c(c("MASS","cluster",
                              "flowWorkspace","flowCore",
                              "hilbertSimilarity","entropy",
                              "Radviz","cytofan",
                              "lme4","emmeans",
                              "FlowSOM","Rtsne",
                              "yaml","tidyverse")))
```

# Radviz

To visualize single cell data associated with this experiment we use Radviz projections ([Abraham 2016](https://doi.org/10.1002/cyto.b.21499)). First we prepare a scaled dataset for projection:

```{targets live.scaled, tar_simple=TRUE}
live.df %>%
    mutate(across(all_of(with(markers,
                              name[type %in% c("phenotypic",
                                               "functional")])),
                  ~ do.L(.x,
                         fun=function(x) quantile(x,
                                                  c(0,0.975))))) %>% 
    select(all_of(c("name",
                    with(markers,
                         name[type %in% c("phenotypic",
                                          "functional")])))) %>% 
    rename_with(~ with(markers,
                       desc[match(.x,name)]),
                !name)
```

Next we Compute a similarity matrix and optimize channel orders for phenotypic and functional channels independently:

```{targets radviz-optim}
list(
    tar_target(live.sim,{
        live.scaled %>% 
            select(!name) %>% 
            as.matrix(.) %>% 
            cosine(.)
    }),
    tar_target(pheno.S,{
        .pheno.S <- make.S(with(markers,
                                desc[type=="phenotypic"]))
        .pheno.optim <- lapply(seq(1,10),function(i) {
            cat('Run #',i,'\n')
            .optim.cur.cells <- do.optimRadviz(.pheno.S,live.sim,top=75)
            return(list(best=in.da(make.S(get.optim(.optim.cur.cells)),live.sim),
                        springs=get.optim(.optim.cur.cells)))
        })
        .pheno.optim.best <- lapply(.pheno.optim,function(x) x$best)
        make.S(.pheno.optim[[which.max(.pheno.optim.best)]]$springs)
    }),
    tar_target(func.S,{
        .func.S <- make.S(with(markers,
                               desc[type=="functional"]))
        .func.optim <- lapply(seq(1,10),function(i) {
            cat('Run #',i,'\n')
            .optim.cur.cells <- do.optimRadviz(.func.S,live.sim,top=75)
            return(list(best=in.da(make.S(get.optim(.optim.cur.cells)),live.sim),
                        springs=get.optim(.optim.cur.cells)))
        })
        .func.optim.best <- lapply(.func.optim,function(x) x$best)
        make.S(.func.optim[[which.max(.func.optim.best)]]$springs)
    })
)
```

Finally we create the `Radviz` objects corresponding to each type of channels:

```{targets radviz-proj}
list(
    tar_target(pheno.rv,{
        live.scaled %>% 
            left_join(annots,
                      by = "name") %>% 
            do.radviz(.,
                      pheno.S)
    }),
    tar_target(func.rv,{
        live.scaled %>% 
            left_join(annots,
                      by = "name") %>% 
            do.radviz(.,func.S)
    })
)
```

# Render the `_explore` report

At this stage we define an `_explore` report that will become part of the pipeline and will be automatically updated as required, to visualize the results of the QC and the `Radviz` projections:

```{targets explore-report}
tarchetypes::tar_render(explore,
                        "reports/_explore.Rmd",
                        params = list(store = tar_config_get("store")))
```

# Run the pipeline

```{r}
tar_make()
```
