---
title: "Cytometry Workflow 5 - Differential Analysis"
output: 
    html_document:
        df_print: paged
        code_folding: show
        toc: yes
        toc_float: yes
        toc_depth: 4
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r}
library(targets)
library(tarchetypes)
```

# Globals

We first define some global options/functions common to all targets.

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c(c("MASS","cluster",
                              "flowWorkspace","flowCore",
                              "hilbertSimilarity","entropy",
                              "Radviz","cytofan",
                              "lme4","emmeans",
                              "FlowSOM",
                              "yaml","tidyverse")))
source("scripts/900_functions.R")
```

# Differential analysis

We run differential analysis using a **G**eneralized **L**inear **M**ixed **M**odel framework with a negative binomial distribution to identify groups of cells that are significantly enriched or depleted in the experiment.

To run statistical modeling we need to define:
  - The smallest (log~2~) fold change we will consider significant (default: 1, or 2-fold fold change)
  - The limit of significance after multiple testing correction (default: 5%)
  - the model used to assess the effect of experimental parameters on population size; the `offset` term defines how changes in population size will be estimated, relative to sample (the default) or relative to the parent population
  - The contrasts that correspond to specific scientific questions

The pipeline will automatically correct for multiple testing at the contrast level (`tukey` method) and at the population level (`fdr` method).

```{targets params-analysis}
list(
    tar_target(params_fold_lim,1),
    tar_target(params_pval_lim,0.05),
    tar_target(live_model_total,
               NULL # Count ~ Stimulation + (1|Individual) + offset(logTotal)
               ),
    tar_target(params_live_contrasts,{
        NULL # list("pos to neg" = c(-1,1))
    }),
    # the following target should not be edited
    tar_target(live_model_specs, {
        .specs <- labels(terms(live_model_total))
        .specs <- .specs[-c(grep("\\|", .specs), grep("\\:", .specs))]
        as.formula(paste("~",
                         paste(.specs,
                               collapse = "+")))
    })
)
```

## Population (manual gates)

Let's start by extracting population sizes:

```{targets pops.df, tar_simple=TRUE}
.gs <- load_gs(params_path)
recompute(.gs) # just in case
.counts <- gs_pop_get_count_fast(.gs)
.total <- .counts %>% 
    filter(Population==params_population) %>% 
    select(name,Total = Count)
annots %>%
    left_join(.counts,
              by = "name") %>%
    left_join(.total,
              by = "name") %>%
    filter(str_detect(Population,fixed(paste0(params_population,"/")))) %>% 
    mutate(logTotal = log(Total),
           logParent = log(ParentCount))
```

We keep both the Parent size and the Total size so that analysis can be run relative to Parent population or over all cells per file. In the next step we fit the models for each population:

```{targets live.pop.models, tar_simple=TRUE}
pops.df %>% 
    nest(data = -Population) %>% 
    mutate(model = map(data,
                       ~ glmer.nb(live_model_total,
                                  data=.,
                                  control = glmerControl(optCtrl = list(maxfun = 25000)))),
           estimate = map2(model,
                           data,
                           ~ emmeans(.x,
                                     specs = live_model_specs,
                                     data = .y)),
           `N cells` = map_dbl(data,~sum(.$Count)),
           `Percent cells` = 100*`N cells`/sum(`N cells`))
```

Next we extract the contrasts:

```{targets live.pop.contrasts, tar_simple=TRUE}
live.pop.models %>% 
    mutate(contrast = map(estimate,~ contrast(.,
                                              params_live_contrasts,
                                              adjust = "tukey"))) %>% 
    select(Population, contrast, 
           `Percent cells`, `N cells`) %>% 
    mutate(contrast = map(contrast,~ as_tibble(.))) %>% 
    unnest(contrast) %>% 
    mutate(contrast = factor(contrast,
                             levels = names(params_live_contrasts)),
           p.adjustFDR = p.adjust(p.value,
                                  method = "fdr"),
           log.p.value = -log10(p.adjustFDR))
```

Finally we prepare a basic volcano plot to visualize the results:

```{targets live.pop.volcano, tar_simple=TRUE}
live.pop.contrasts %>% 
    mutate(PopName = str_replace(Population,
                                 fixed(params_population),
                                 "")) %>% 
    ggplot(aes(x = estimate,
               y = log.p.value))+
    geom_label(aes(label = PopName))+
    geom_vline(xintercept = c(-1,1),
    geom_vline(xintercept = c(-params_fold_lim,
                              params_fold_lim),
               col = 2, lty = 2)+
    geom_hline(yintercept = -log10(0.05),
    geom_hline(yintercept = -log10(params_pval_lim),
               col = 2, lty = 2)
```

## Metaclusters

Let's start by extracting population sizes:

```{targets metas.df, tar_simple=TRUE}
live.df %>% 
    mutate(Population = factor(live.fsom.metaclusters)) %>% 
    count(name, Population,
          name = "Count") %>% 
    group_by(name) %>% 
    mutate(Total = sum(Count),
           logTotal = log(Total)) %>% 
    ungroup() %>% 
    left_join(annots,
              by = "name")
```

Only the Total size makes sense here, since metaclusters are not nested. In the next step we fit the models for each population:

```{targets live.meta.models, tar_simple=TRUE}
metas.df %>% 
    nest(data = -Population) %>% 
    mutate(model = map(data,
                       ~ glmer.nb(live_model_total,
                                  data=.,
                                  control = glmerControl(optCtrl = list(maxfun = 25000)))),
           estimate = map2(model,
                           data,
                           ~ emmeans(.x,
                                     specs = live_model_specs,
                                     data = .y)),
           `N cells` = map_dbl(data,~sum(.$Count)),
           `Percent cells` = 100*`N cells`/sum(`N cells`))
```

Next we extract the contrasts:

```{targets live.meta.contrasts, tar_simple=TRUE}
live.meta.models %>% 
    mutate(contrast = map(estimate,~ contrast(.,
                                              params_live_contrasts,
                                              adjust = "tukey"))) %>% 
    select(Population, contrast, 
           `Percent cells`, `N cells`) %>% 
    mutate(contrast = map(contrast,~ as_tibble(.))) %>% 
    unnest(contrast) %>% 
    mutate(contrast = factor(contrast,
                             levels = names(params_live_contrasts)),
           p.adjustFDR = p.adjust(p.value,
                                  method = "fdr"),
           log.p.value = -log10(p.adjustFDR))
```

Finally we prepare a basic volcano plot to visualize the results:

```{targets live.meta.volcano, tar_simple=TRUE}
live.meta.contrasts %>% 
    ggplot(aes(x = estimate,
               y = log.p.value))+
    geom_label(aes(label = Population))+
    geom_vline(xintercept = c(-1,1),
    geom_vline(xintercept = c(-params_fold_lim,
                              params_fold_lim),
               col = 2, lty = 2)+
    geom_hline(yintercept = -log10(0.05),
    geom_hline(yintercept = -log10(params_pval_lim),
               col = 2, lty = 2)
```

# Render the `_analysis` report

We will visualize the results and run more exploration in the `_analysis` report:

```{targets analysis-report}
tarchetypes::tar_render(analysis,
                        "reports/_analysis.Rmd",
                        params = list(store = tar_config_get("store")))
```

# Run the pipeline

```{r run}
tar_make()
```
