---
title: "Cytometry Workflow 7 - Supervized Learning"
output: 
    html_document:
        df_print: paged
        code_folding: show
        toc: yes
        toc_float: yes
        toc_depth: 4
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r libraries}
library(targets)
library(tarchetypes)
```

# Globals

We first define some global options/functions common to all targets.

```{targets globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("tidyverse"))
source("scripts/900_functions.R")
```

# Freeviz

To optimize a [Freeviz]() we need the following parameters:

 - the metacluster(s) of interest 
 - the channels to optimize
 - the variable(s) to optimize for
 - the limit on weights for the channels (visualization only)

```{targets params-freeviz}
list(
    tar_target(params_fv_metas,c(4,5)),
    tar_target(params_fv_channels,{
        with(markers,
             desc[type %in% c("phenotypic","functional")])
    }),
    tar_target(params_fv_conditions,
               c("Condition","PatientType")),
    tar_target(anchor.lim,0.3)
)
```

We can now optimize the channels:

```{targets freeviz-optims, tar_interactive=FALSE}
tar_target(live.fv.optims,{
    library(Radviz)
    .mat <- live.scaled %>%
        left_join(annots,
                  by = "name") %>%
        mutate(Population = live.fsom.metaclusters) %>% 
        unite("conditions",
              all_of(params_fv_conditions)) %>% 
        filter(Population==params_fv_metas)
    
    .weights <- .mat %>% 
        count(conditions,
              name = "weight")
    
    if(nrow(.mat)>10000) {
        .n <- min(10000/nrow(.weights),
                  .weights$weight)
        .mat <- .mat %>% 
            slice_sample(n=.n,
                         by = conditions)
    }
    
    .mat %>% 
        select(all_of(params_fv_channels)) %>% 
        as.matrix() %>% 
        do.optimFreeviz(.,
                        classes = .mat$conditions)
},
pattern = map(params_fv_metas),
iteration = "list")
```

For interpretation we will compute the channel weights:

```{targets freeviz-weights, tar_interactive=FALSE}
tar_target(live.fv.weights,{
    tibble(Channel = rownames(live.fv.optims),
           weight = rowSums(live.fv.optims^2)^0.5) %>% 
        mutate(i = dense_rank(weight))
},
pattern = map(params_fv_metas,live.fv.optims),
iteration = "list")
```

And run a Freeviz projection:

```{targets freeviz-projs, tar_interactive=FALSE}
tar_target(live.fvs,{
    library(Radviz)
    live.scaled %>%
        left_join(annots,
                  by = "name") %>%
        mutate(Population = live.fsom.metaclusters) %>% 
        filter(Population==params_fv_metas) %>% 
        do.radviz(.,
                  live.fv.optims)
},
pattern = map(params_fv_metas,live.fv.optims),
iteration = "list")
```

# Render the `_supervized` report

At this stage we define an `_supervized` report that will become part of the pipeline and will be automatically updated as required, to visualize the results of the `Freeviz` training:

```{targets supervized-report, tar_interactive=FALSE}
tarchetypes::tar_render(supervized,
                        "reports/_supervized.Rmd",
                        params = list(store = tar_config_get("store")))
```

# Run the pipeline

```{r run}
tar_make()
```
