---
title: "Visualizing Supervized Analysis"
params:
  store: !r file.path("..","_targets")
output:
  html_notebook:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
---

# loading the packages

```{r setup}
library(targets)

library(knitr)

library(tidyverse)

library(Radviz)

theme_set(theme_light(base_size = 16))
tar_config_set(store = params$store)
```

# loading the data

```{r data}
tar_load(anchor.lim)
tar_load(params_fv_conditions)
tar_load(live.fv.weights)
tar_load(live.fv)
```

# Visualizing the weights

The channel weights correspond to the relevance of a particular channel to the separation between training categories:

```{r}
live.fv.weights %>% 
    ggplot(aes(x = i,
               y = weight))+
    geom_text(aes(label = Channel))+
    geom_hline(yintercept = anchor.lim,
               col = 2,
               lty = 2)
```

The red dotted line corresponds to the `anchor.lim` parameter and will be used to filter out channels that don't contribute to the separation between classes.

# Visualizing the `Freeviz` results

```{r}
plot(
    rescalePlot(
        anchor.filter(live.fv,anchor.lim)))+
    geom_density2d(data = .  %>% 
                       unite("conditions",
                             all_of(params_fv_conditions)) %>% 
                       slice_sample(prop = 0.1,
                                    by = conditions),
                   aes(color = conditions))
```

