---
title: "Visualizing Supervized Analysis"
params:
  store: !r file.path("..","_targets")
output:
  html_notebook:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
---

# loading the packages

```{r setup}
library(targets)

library(knitr)

library(tidyverse)

library(Radviz)

theme_set(theme_light(base_size = 16))
tar_config_set(store = params$store)
```

# loading the data

```{r data}
tar_load(anchor.lim)
tar_load(params_fv_metas)
tar_load(params_fv_conditions)
tar_load(live.fv.weights)
tar_load(live.fvs)
```

To simplify plotting we preprocess the `Freeviz` plots:

```{r freeviz-preprocess}
live.fvs <- lapply(live.fvs,anchor.filter,anchor.lim)
live.fvs <- lapply(live.fvs,function(fv) suppressWarnings(rescalePlot(fv)))
```

# Visualizing the weights {.tabset}

The channel weights correspond to the relevance of a particular channel to the separation between training categories.

The red dotted line corresponds to the `anchor.lim` parameter and will be used to filter out channels that don't contribute to the separation between classes.

```{r weights, fig.width = 6, fig.height = 6, results="asis"}
ksink <- lapply(seq(length(params_fv_metas)), function(imetas) {
    cat("## Metacluster #",params_fv_metas[imetas],"\n\n")
    p <- live.fv.weights[[imetas]] %>% 
        ggplot(aes(x = i,
                   y = weight))+
        geom_text(aes(label = Channel))+
        geom_hline(yintercept = anchor.lim,
                   col = 2,
                   lty = 2)
    print(p)
    cat("\n\n")
})
```

# Visualizing the `Freeviz` results {.tabset}

```{r freevizs, fig.width = 8, fig.height = 6, results="asis"}
ksink <- lapply(seq(length(params_fv_metas)), function(imetas) {
    cat("## Metacluster #",params_fv_metas[imetas],"\n\n")
    p <- plot(live.fvs[[imetas]])+
        geom_density2d(data = .  %>% 
                           unite("conditions",
                                 all_of(params_fv_conditions)) %>% 
                           slice_sample(n = 5000,
                                        by = conditions,
                                        replace = TRUE),
                       aes(color = conditions))
    print(p)
    cat("\n\n")
})
```

